---
title: "Accessibilty Lecture"
author: "Carole Voulgaris"
date: "7/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tigris)
library(osmdata)
library(opentripplanner)
library(tidyverse)
library(leaflet)
library(sf)
```

## What is accessibilty?

The ease with which a person can fully participate in the economic, civic, and social life of their society. Accessibility can be described in terms of three components:

* Connectivity
* Mobility
* Proximity

**Question to consider:** Is travel a derived demand?

## Connectivity

The ability to participate in activities without traveling.

## Mobility 

The ability to move through space. Can be measured as:

* Distance you can travel per unit of time (speed)
* Distance you can travel per dollar
* Area within 30 minutes
* (all of these are mode-dependent)

```{r, echo = FALSE}
Ballard <- c(lon = -122.387519, lat = 47.668740)
Downtown <- c(lon = -122.339258, lat = 47.611858)
Kirkland <- c(lon = -122.168838, lat = 47.723188)
box <- st_bbox(King_CO)

path_data <- file.path(getwd(), "OTP")
path_otp <- paste(path_data,"otp.jar",sep = "/")

otp_setup(otp = path_otp, dir = path_data, memory = 1024)

# Connect to opentripplanner
otpcon <- otp_connect()
```

```{r}

Downtown_drive_30 <- otp_isochrone(otpcon = otpcon, 
                              fromPlace = Downtown,
                              mode = "CAR",
                              cutoffSec = 1800) %>%
  st_transform(crs = "+proj=longlat +datum=WGS84")

Downtown_walk_30 <- otp_isochrone(otpcon = otpcon, 
                              fromPlace = Downtown,
                              mode = "WALK",
                              maxWalkDistance = 6000,
                              cutoffSec = 1800) %>%
  st_transform(crs = "+proj=longlat +datum=WGS84")

Downtown_transit_30 <- otp_isochrone(otpcon = otpcon, 
                              fromPlace = Downtown,
                              mode = c("WALK", "TRANSIT"),
                              maxWalkDistance = 6000,
                              cutoffSec = 1800) %>%
  st_transform(crs = "+proj=longlat +datum=WGS84")


isochrones = leaflet(Downtown_drive_30) %>%
  addPolygons(color = NA, fillColor = "red") %>%
  #addPolygons(data = Kirkland_transit_30, 
  #            color = NA, fillColor = "green") %>%
  #addPolygons(data = Kirkland_walk_30, 
  #            color = NA, fillColor = "blue") %>%
  #addCircleMarkers(lng = Kirkland["lon"],
  #                 lat = Kirkland["lat"],
  #                 radius = 3,
  #                 color = NA,
  #                 fillColor = "Red",
  #                 fillOpacity = 1) %>%
  addProviderTiles(providers$Stamen.Toner)
isochrones

st_write(Downtown_drive_30, "dwntwn_drive.shp")
st_write(Downtown_walk_30, "dwntwn_walk.shp")
st_write(Downtown_transit_30, "dwntwn_transit.shp")

```

## Proximity metrics

The amount of space between a location and opportunities to participate in activities. Can be measured as:

* Distance required to reach a destination
* Number of destinations within 10 miles

## What do we mean by distance?

* Euclidean distance
* Network distance (mode-dependent if some links are restricted by mode)

## Accessibilty metrics

* Number of opportunities within 30 minutes
* Total number of opportunities, weighted based on how long it takes to reach them.


```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
