---
title: "Accessibilty Lecture"
author: "Carole Voulgaris"
date: "7/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tigris)
library(osmdata)
library(opentripplanner)
library(tidyverse)
library(leaflet)
library(sf)
```

## What is accessibilty?

The ease with which a person can fully participate in the economic, civic, and social life of their society. Accessibility can be described in terms of three components:

* Connectivity
* Mobility
* Proximity

**Question to consider:** Is travel a derived demand?

## Connectivity

The ability to participate in activities without traveling.

## Mobility 

The ability to move through space. Can be measured as:

* Distance you can travel per unit of time (speed)
* Area within 30 minutes
* Distance you can travel for a cost of $20
* (all of these are mode-dependent)

```{r, echo = FALSE}
Site <- c(lon = -71.149450, lat = 42.208314)

state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"

East_Mass <- counties(state = "MA", class = "sf") %>%
  filter(GEOID == "25017" |
         GEOID == "25021" |
         GEOID == "25025")

box <- st_bbox(East_Mass)
```

```{r}

roads <- opq(bbox = box, timeout = 60) %>%
  add_osm_feature(key = 'highway') %>% 
  osmdata_xml(file = 'OTP/graphs/default/MA_Roads.osm')

path_data <- file.path(getwd(), "OTP")
path_otp <- paste(path_data,"otp.jar",sep = "/")
#path_otp <- otp_dl_jar(path_data)

# You only need to run this line once.

log1 <- otp_build_graph(otp = path_otp, dir = path_data, memory = 1024) 
```

```{r}

log2 <- otp_setup(otp = path_otp, dir = path_data)

# Connect to opentripplanner
otpcon <- otp_connect()
```

```{r}

drive_30_min <- otp_isochrone(otpcon = otpcon, 
                              fromPlace = Site,
                              mode = "CAR",
                              cutoffSec = 1800) %>%
  st_transform(crs = state_plane)

isochrones = leaflet(drive_30_min,
                      options = leafletOptions(minZoom = 14,
                                               maxZoom = 18)) %>%
  addProviderTiles(providers$Stamen.Toner) %>%
  setView(lat = Site["lat"], 
          lng = Site["lon"],
          zoom = 14) %>%
  addPolygons(fill = FALSE, color = "black")


```

## Proximity metrics

The amount of space between a location and opportunities to participate in activities. Can be measured as:

* Distance required to reach a destination
* Number of destinations within 10 miles

## What do we mean by distance?

* Euclidean distance
* Network distance (mode-dependent if some links are restricted by mode)

## Accessibilty metrics

* Number of opportunities within 30 minutes
* Total number of opportunities, weighted based on how long it takes to reach them.


```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
